Exploratory Analysis on CDC Teen Vaccination Surveys

### 2024-05-03 code/analysis-code/analysis2.qmd
Final edits and cleaning

---
title: "Data Analysis and Modelling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence

@@

### Conclusion
our analysis aimed to address collinearity issues in our teen vaccination surveys dataset through dimension reduction techniques like PCA and regularization methods such as LASSO regression. LASSO regression effectively selected important predictors while simplifying the model. Subsequent cross-validation demonstrated consistent model performance, with an average accuracy of 79.4% and robust precision and recall metrics.

>>>>

<<<<<<< Updated upstream
---
title: "Data Analysis and Modelling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence

@@

### Conclusion
=======
---
title: "Data Analysis and Modelling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence

### Conclusion
>>>>>>> Stashed changes
our analysis aimed to address collinearity issues in our teen vaccination surveys dataset through dimension reduction techniques like PCA and regularization methods such as LASSO regression. LASSO regression effectively selected important predictors while simplifying the model. Subsequent cross-validation demonstrated consistent model performance, with an average accuracy of 79.4% and robust precision and recall metrics.

### 2024-05-03 code/analysis-code/analysis4.qmd
Final edits and cleaning

Despite this, I will see if the important predictors are in agreement with our LASSO model using the vip package.
```{r}
# Finding importance of variables chosen
en_vars <- en_tune_fit %>%
  extract_fit_parsnip() %>% # use parsnip to extract to fit
  vip::vip() #use vip for visualizing the most important variables

print(en_vars)
```
The elastic net model produced the same important variables as the previous lasso model.

>>>>

We will examine the coefficients of the model to determine the associations of the variables with the outcome.
```{r}
# Extract and tidy model parameters
en_tune_coef <- en_tune_fit %>%
  extract_fit_parsnip() %>%
  tidy()

# Filter out predictors with non-zero coefficients
en_nonzero_coef <- en_tune_coef %>%
  filter(estimate != 0)
en_nonzero_coef
```

Despite this, I will see if the important predictors are in agreement with our LASSO model using the vip package.

The elastic net model produced the same important variables as the previous lasso model. 
```{r}
# Finding importance of variables chosen
en_vars <- en_tune_fit %>%
  extract_fit_parsnip() %>% # use parsnip to extract to fit
  vip::vip() #use vip for visualizing the most important variables

# Modify the plot to set the color of the bars
en_vars_plot <- en_vars+
  aes(fill = "skyblue")+
  labs(title = "Top Elastic Net Predictors")

# Create a data frame with original and translated variable names
variable_translations <- data.frame(
  Original_Variable = rev(c("STATE_RHODE.ISLAND", "FACILITY", 
                        "STATE_VERMONT", "STATE_MISSISSIPPI", "STATE_NORTH.DAKOTA", "STATE_IOWA", 
                        "STATE_HAWAII", "RENT_OWN_DON.T.KNOW", "STATE_NEBRASKA", 
                        "INS_BREAK_I_CURRENTLY.UNINSURED.AND.NEVER.INSURED.SINCE.AGE.11")),
  Translated_Variable = rev(c("Rhode Island", "All Std School Teen Clinics or Other Facilities", 
                          "Vermont", "Mississippi", "North Dakota", "Iowa", "Hawaii", 
                          "Rent/Own Status (Don't Know)", "Nebraska", 
                          "Uninsured"))
)

# Add translations to the plot
en_plot_with_translations <- en_vars_plot +
  scale_fill_manual(values = c("skyblue")) +  # Set fill color manually
  scale_x_discrete(labels = variable_translations$Translated_Variable) +  # Translate x-axis labels
  labs(title = "Top Predictors",
                x = "Variable")+  # Add title and update x-axis label
  scale_y_continuous(labels = function(x) strwrap(x, width = 10))

# Display the plot with translated variable names
print(en_plot_with_translations)

# Save the plot as a PNG file
path = here("results", "figures", "top_en_predictors.png")
ggsave(filename = path, plot=en_plot_with_translations) 
```

### 2024-05-03 code/eda-code/EDA-code.qmd
Final edits and cleaning

```{r, warnings=FALSE, message=FALSE}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(knitr)
```

@@

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results","tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

>>>>

```{r, warnings=FALSE, message=FALSE}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(knitr)
library(gtsummary)
```

@@

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file

summarytable_file = here("results","tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```
I will create a table with gt summary to display in the manuscript.
```{r}
table1 <- mydata %>%
  select(AGE, SEX, RACEETHK, INS_STAT2_I, INCPOV1, P_UTDHPV) %>% # select the variables to include in table
  mutate(SEX = factor(SEX, # change the factor labels using dpylr
                      levels = c("MALE", "FEMALE"),
                      labels = c("Male", "Female")),
         RACEETHK = factor(RACEETHK, 
                           levels = c("NON-HISPANIC WHITE ONLY", 
                                      "NON-HISPANIC BLACK ONLY", 
                                      "NON-HISPANIC ASIAN ONLY",
                                      "HISPANIC",
                                      "NON-HISPANIC OTHER + MULTIPLE RACE"),
                           labels = c("White", "Black", "Asian", "Hispanic", "Multiracial")),
         INS_STAT2_I = factor(INS_STAT2_I,
                              levels = c("PRIVATE INSURANCE ONLY", 
                                         "ANY MEDICAID", 
                                         "UNINSURED",
                                         "OTHER INSURANCE (CHIP, IHS, MILITARY, OR OTHER, ALONE OR IN COMB. WITH PRIVATE INSURANCE)"),
                              labels = c("Private", "Medicaid", "Uninsured", "Other insurance")),
         INCPOV1 = factor(INCPOV1,
                          levels = c("BELOW POVERTY", 
                                     "ABOVE POVERTY > $75K", 
                                     "ABOVE POVERTY <= $75K", 
                                     "UNKNOWN"),
                          labels = c("Below Poverty", "Above Poverty > $75K", "Above poverty <= $75K", "Unknown")),
         P_UTDHPV = factor(P_UTDHPV,
                           levels = c("UTD", "NOT UTD"),
                           labels = c("Up-to-date", "Not Up-to-date")))%>%
  tbl_summary(by = P_UTDHPV, # group by HPV vaccination status
      statistic = list( 
      all_continuous() ~ "{mean} ({sd})", # select the mean and standard deviation of continuous variables
      all_categorical() ~ "{n} ({p}%)"),# select the number and percentage of categorical variables
    label = list(
      AGE ~ "Age (years)", # change the names of the variables to be more readable
      SEX ~ "Sex",   
      RACEETHK ~ "Ethnicity", 
      INS_STAT2_I ~ "Insurance Status",      
      INCPOV1 ~ "Poverty Status"), missing = "no") %>%
    bold_labels() %>%
  modify_caption("Table 1. Study sample of Teenagers aged 13-17 who received the NIS Teen survey in the U.S.")

table1

# Save the summary table
saveRDS(table1, file = here("results/tables/summary_table.rds"))
```
